---
title: "IOR Control Experiment Data Analysis" 
author: "P. Zacher" 
date: "2023-11-01" 
output: html_document 
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages

```{r, warning=FALSE, message=FALSE}
library(psych) # for stats
library(psychReport) # for reproducible reports
library(readr) # for importing data
library(Hmisc) # for plotting
library(Rmisc) # for standard error calc
library(readxl) # for reading excel files
library(tidyverse) # packages for data science
library(plyr)
library(ez) # for stats
library(statmod)
library(ggplot2) # plotting
library(knitr) # for formatting tables
```

### Import Data

The data contains observations of all trials for each participant who
completed the task. Variables of interest are the trial number
(`trialNum`), duration the participant had to orient, engage, and
disengage attention (`ISIadj`), the number of attention shifts required
to attend to the target digit (`targetIndex`), whether the participant
answered correctly (`correct`), and the number of staircase reversals
completed for each trial type (`nReversals`).

```{r}
behavioral_data <- read.csv(file = "visual_shifting_ior_control_data.csv")
```

## Data Cleaning

Minimal data cleaning and transformation completed here as most was
accomplished prior to importing the data. We are primarily interested in
performance at threshold, so we will filter all trials that are not at
threshold (`nReversal` != 0). A small number of reversals may not occur
at threshold, but they likely do not impact our performance measure when
included in the data set. We will also convert attention shifting times,
`ISIadj`, to milliseconds (ms). Lastly, we will convert `targetIndex`
and `p` to factor because they are categorical variables.

```{r, results='asis'}
behavioral_data_clean <- behavioral_data %>% filter(nReversal != 0)
behavioral_data_clean$ISIms <- behavioral_data_clean$ISIadj*1000
behavioral_data_clean$targetIndex <- as.factor(behavioral_data_clean$targetIndex)
behavioral_data_clean$p <- as.factor(behavioral_data_clean$p)
```

### Descriptive Statistics

Create a table with summary statistics (e.g., mean, standard deviation,
and within-subject standard error). We will use these values to visually
identify potential outliers.

```{r echo = FALSE, results = 'asis'}
# Mean shifting speed for each participant and target index
behavioral_data_summ_p <- behavioral_data_clean %>%
  dplyr::group_by(p, targetIndex) %>%
  dplyr::summarize(meanISIms = mean(ISIms)) %>% 
  ungroup()

# Use `summarySEwithin` function from the Rmisc package to calculate within-
# subject standard error.
behavioral_data_summ_se <- summarySEwithin(behavioral_data_summ_p,
                                           measurevar = "meanISIms",
                                           withinvars = "targetIndex",
                                           idvar = "p")

# Add columns calculating overall mean +/- 3 standard deviations as this
# criteria will be used if we need to remove outliers.
behavioral_data_summ_se <- mutate(behavioral_data_summ_se, 
                                  "mean_plus_3sd" = meanISIms + 3*sd,
                                  "mean_minus_3sd" = meanISIms - 3*sd)

# Create a nicely formatted table with summary statistics
kable(behavioral_data_summ_se,
      col.names = c("Target Index", "N", "Duration (ms)", "SD", "SE", "CI", "M + 3SD", "M - 3SD"), 
      align = "c", 
      caption = "Shifting Speed Summary Statistics",
      digits = 2)

```

```{r}
# Boxplot summarizing performance and identifying potential outliers
ggplot(behavioral_data_summ_p,
       aes(x = targetIndex,
           y = meanISIms,
           fill = targetIndex)) +
  geom_boxplot(outlier.color = "red") +
  stat_summary(fun.y = mean, 
               geom = "point", 
               shape = 21, 
               size = 2,
               color = "black",
               fill = "black") + # Add means
  geom_point(data = behavioral_data_summ_se, # Add upper outlier cutoff (Mean + 3*SD)
             aes(x = targetIndex,
                 y = mean_plus_3sd),
             color = "black",
             fill = "gray",
             shape = 24) + 
  geom_point(data = behavioral_data_summ_se, # Add lower outlier cutoff (Mean - 3*SD)
             aes(x = targetIndex,
                 y = mean_minus_3sd),
             color = "black",
             fill = "gray",
             shape = 25) +
    scale_x_discrete(labels=c("0" = "0", # Labels for dependent variables
                            "1" = "1",
                            "2" = "2",
                            "3" = "3",
                            "7" = "7", 
                            "8" = "8",
                            "9" = "9",
                            "99" = "Catch Trials")) +
  xlab("Number of Shifts") + 
  ylab("Mean Shifting Speed") + 
  labs(caption = "Lines within each box represent the median. Black points represent means. Red points represent\n participant means that fall outside of the 1.5*IQR (IQR = distance between 1st and 3rd quantile)\n threshold. Triangles represent the threshold for outlier performance (Mean +/- 3*SD).") +
  theme_light() +
  theme(legend.position = "none")
```

### Identify and remove outliers

We will use a conservative criterion for outlier removal. We can expect 99.72%
of all observations to be contained within 3 standard deviations from the mean.
As such, any participants whose performance is greater than this threshold is
likely not to be representative of the population. Therefore, we will remove 
those participants whose performance in any one condition was greater than 3 
standard deviations from the mean.

```{r}
# Join `behavioral_data_summ_p` and `behavioral_data_summ_se`.
behavioral_data_summ <- left_join(behavioral_data_summ_p, 
                                  behavioral_data_summ_se, 
                                  by = 'targetIndex',
                                  suffix = c('_p', '_all_p'))

# Add a column that specifies whether the participant's values for a given
# condition surpass the outlier threshold.
behavioral_data_summ <- behavioral_data_summ %>% 
  mutate(outlier = meanISIms_p > mean_plus_3sd | meanISIms_p < mean_minus_3sd)

# Create a vector containing the participant IDs contributing outliers.
outlier_p <- unique(behavioral_data_summ$p[behavioral_data_summ$outlier == TRUE])

# Filter the participants contributing outliers.
behavioral_data_summ_outlier_removed <- behavioral_data_summ %>% 
  filter(!p %in% outlier_p)

# Boxplot summarizing performance without outliers
ggplot(behavioral_data_summ_outlier_removed,
       aes(x = targetIndex,
           y = meanISIms_p,
           fill = targetIndex)) +
  geom_boxplot(outlier.color = "red") +
  stat_summary(fun.y = mean, 
               geom = "point", 
               shape = 21, 
               size = 2,
               color = "black",
               fill = "black") + # Add means
  geom_point(aes(x = targetIndex, # Add upper outlier cutoff (Mean + 3*SD)
                 y = mean_plus_3sd),
             color = "black",
             fill = "gray",
             shape = 24) + 
  geom_point(aes(x = targetIndex, # Add lower outlier cutoff (Mean - 3*SD)
                 y = mean_minus_3sd),
             color = "black",
             fill = "gray",
             shape = 25) +
    scale_x_discrete(labels=c("0" = "0", 
                            "1" = "1",
                            "2" = "2",
                            "3" = "3",
                            "7" = "7", 
                            "8" = "8",
                            "9" = "9",
                            "99" = "Catch Trials")) +
  xlab("Number of Shifts") + 
  ylab("Mean Shifting Speed") + 
  labs(caption = "Lines within each box represent the median. Black points represent means. Red points represent\n participant means that fall outside of the 1.5*IQR (IQR = distance between 1st and 3rd quantile)\n threshold. Triangles represent the threshold for outlier performance (Mean +/- 3*SD).") +
  theme_light() +
  theme(legend.position = "none")

```

### Figures

```{r}
# Bar plot with points for mean performance for each participant and error bars
# representing the standard error corrected for within-subjects effects.
ggplot(behavioral_data_summ_se,
       aes(x = targetIndex,
           y = meanISIms,
           fill = targetIndex)) +
  geom_col(width = 0.75, color = "black") + 
  geom_errorbar(aes(y = meanISIms, 
                    ymin = meanISIms - se,
                    ymax = meanISIms + se,
                    width = 0.05)) +
  geom_point(data = behavioral_data_summ_p, 
             aes(x = targetIndex,
                 y = meanISIms),
             color = "black", 
             fill = "black", 
             shape = 21, 
             size = 1, 
             alpha = 0.3) + 
  scale_x_discrete(labels=c("0" = "0", 
                            "1" = "1",
                            "2" = "2",
                            "3" = "3",
                            "7" = "7", 
                            "8" = "8",
                            "9" = "9",
                            "99" = "Catch")) +
  xlab("Number of Shifts") + 
  ylab("Mean Shifting Speed") + 
  labs(caption = "Bar plots represent the mean performance per condition. Points represent participant means, error bars represent 
       the within-subject standard error (Morey, 2008), calculated with the Rmisc package (Hope, 2013).") +
  theme_light() +
  theme(legend.position = "none")

```

### ANOVA

```{r}
# Convert `participant` and `target index` cols to factor
behavioral_data_clean <- behavioral_data_clean %>%
  mutate(p = as.factor(p), targetIndex = as.factor(targetIndex))

# Compute 1-way ANOVA where `target index` is the within-subjects factor
behavioral_data_clean_ANOVA <- ezANOVA(
  behavioral_data_clean,
  dv = .(ISIms),
  wid = .(p),
  within = .(targetIndex),
  return_aov = TRUE,
  detailed = TRUE
)

# Change effect size to partial eta squared (pes)
behavioral_data_clean_ANOVA <- aovEffectSize(behavioral_data_clean_ANOVA, 
                                             effectSize = "pes")

# Create a nicely formatted table for results
aovDispTable(behavioral_data_clean_ANOVA)

```

## Post-Hoc Tests - Tukey's HSD

```{r}
library(nlme) # Add to library if we end up using this approach.
library(multcomp) # For glht pos-hoc tests

# Create a new model using lme.
planned_comparisons <- lme(ISIms ~ targetIndex, 
                           random = ~1|p/targetIndex, # Random effects resulting from participants 
                           data = behavioral_data_clean, 
                           method = "ML") # Maximum likelihood

# Post-hoc tests (Tukey).
post_hocs <- glht(planned_comparisons, linfct = mcp(targetIndex = "Tukey"))
summary(post_hocs)

```
