---
title: "IOR Control Experiment Data Analysis" 
author: "P. Zacher" 
date: "2023-11-01" 
output: html_document 
editor_options: 
  markdown: 
    wrap: 72
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages
```{r, warning=FALSE, message=FALSE}
library(psych) # for stats
library(readr) # for importing data
library(Hmisc) # for plotting
library(Rmisc) # for standard error calc
library(readxl) # for reading excel files
library(tidyverse) # packages for data science
library(plyr)
library(ez) # for stats
library(statmod)
library(ggplot2) # plotting
library(knitr) # creating nice tables
```

### Import Data 
The data contains observations for each participant who completed the task and
for each trial. Variables of interest are the trial number (`trialNum`),
duration the participant had to orient, engage, and disengage attention
(`ISIadj`), the number of attention shifts required to attend to the target
digit (`targetIndex`), whether the participant answered correctly (`correct`),
and the number of staircase reversals completed for each trial type
(`nReversals`).
```{r}
behavioral_data <- read.csv(file = "visual_shifting_ior_control_data.csv")
```

## Data Cleaning
Filter trials not at threshold (nReversal != 0). 
```{r, results='asis'}
behavioral_data_clean <- behavioral_data %>% filter(nReversal != 0)
behavioral_data_clean$ISIms <- behavioral_data_clean$ISIadj*1000
behavioral_data_clean$targetIndex <- as.factor(behavioral_data_clean$targetIndex)
```

### Figures
```{r}
# Mean shifting speed for each participant and target index
behavioral_data_summ_p <- behavioral_data_clean %>%
  dplyr::group_by(p, targetIndex) %>%
  dplyr::summarize(meanISIms = mean(ISIms))

# Use `summarySEwithin` function from the Rmisc package to calculate within-
# subject standard error.
behavioral_data_summ_se <- summarySEwithin(behavioral_data_summ_p,
                                           measurevar = "meanISIms",
                                           withinvars = "targetIndex",
                                           idvar = "p")

# Bar plot with points for mean performance for each participant and error bars
# representing the standard error corrected for within-subjects effects.
ggplot(behavioral_data_summ_se,
       aes(x = targetIndex,
           y = meanISIms,
           fill = targetIndex)) +
  geom_col(width = 0.75, color = "black") + 
  geom_errorbar(aes(y = meanISIms, 
                    ymin = meanISIms - se,
                    ymax = meanISIms + se,
                    width = 0.05)) +
  geom_point(data = behavioral_data_summ_p, 
             aes(x = targetIndex,
                 y = meanISIms),
             color = "black", 
             fill = "white", 
             shape = 21, 
             size = 1, 
             alpha = 0.75) + 
  xlab("Number of Shifts") + 
  ylab("Shifting Speed") + 
  labs(caption = "Bar plots represent the mean performance per condition. Points represent participant means, error bars represent 
       the within-subject standard error (Morey, 2008), calculated with the Rmisc package (Hope, 2013).") +
  theme_light() +
  theme(legend.position = "none")

```
