---
title: "IOR Control Experiment Data Analysis" 
author: "P. Zacher" 
date: "2023-11-01" 
output: html_document 
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages

```{r, warning=FALSE, message=FALSE}
library(psych) # for stats
library(psychReport) # for reproducible reports
library(readr) # for importing data
library(Hmisc) # for plotting
library(Rmisc) # for standard error calc
library(readxl) # for reading excel files
library(tidyverse) # packages for data science
library(plyr)
library(ez) # for stats
library(statmod)
library(ggplot2) # plotting
library(knitr) # for formatting tables
```

### Import Data

The data contains observations for each participant who completed the
task and for each trial. Variables of interest are the trial number
(`trialNum`), duration the participant had to orient, engage, and
disengage attention (`ISIadj`), the number of attention shifts required
to attend to the target digit (`targetIndex`), whether the participant
answered correctly (`correct`), and the number of staircase reversals
completed for each trial type (`nReversals`).

```{r}
behavioral_data <- read.csv(file = "visual_shifting_ior_control_data.csv")
```

## Data Cleaning

Filter trials not at threshold (nReversal != 0).

```{r, results='asis'}
behavioral_data_clean <- behavioral_data %>% filter(nReversal != 0)
behavioral_data_clean$ISIms <- behavioral_data_clean$ISIadj*1000
behavioral_data_clean$targetIndex <- as.factor(behavioral_data_clean$targetIndex)
```

### Figures

```{r echo = FALSE, results = 'asis'}
# Mean shifting speed for each participant and target index
behavioral_data_summ_p <- behavioral_data_clean %>%
  dplyr::group_by(p, targetIndex) %>%
  dplyr::summarize(meanISIms = mean(ISIms))

# Use `summarySEwithin` function from the Rmisc package to calculate within-
# subject standard error.
behavioral_data_summ_se <- summarySEwithin(behavioral_data_summ_p,
                                           measurevar = "meanISIms",
                                           withinvars = "targetIndex",
                                           idvar = "p")

# Bar plot with points for mean performance for each participant and error bars
# representing the standard error corrected for within-subjects effects.
ggplot(behavioral_data_summ_se,
       aes(x = targetIndex,
           y = meanISIms,
           fill = targetIndex)) +
  geom_col(width = 0.75, color = "black") + 
  geom_errorbar(aes(y = meanISIms, 
                    ymin = meanISIms - se,
                    ymax = meanISIms + se,
                    width = 0.05)) +
  geom_point(data = behavioral_data_summ_p, 
             aes(x = targetIndex,
                 y = meanISIms),
             color = "black", 
             fill = "white", 
             shape = 21, 
             size = 1, 
             alpha = 0.75) + 
  xlab("Number of Shifts") + 
  ylab("Shifting Speed") + 
  labs(caption = "Bar plots represent the mean performance per condition. Points represent participant means, error bars represent 
       the within-subject standard error (Morey, 2008), calculated with the Rmisc package (Hope, 2013).") +
  theme_light() +
  theme(legend.position = "none")

kable(behavioral_data_summ_se,
      col.names = c("Target Index", "N", "Duration (ms)", "SD", "SE", "CI"), 
      align = "c", 
      caption = "Shifting Speed Summary Statistics",
      digits = 2)

```

### ANOVA

```{r}
# Convert `participant` and `target index` cols to factor
behavioral_data_clean <- behavioral_data_clean %>%
  mutate(p = as.factor(p), targetIndex = as.factor(targetIndex))

# Compute 1-way ANOVA where `target index` is the within-subjects factor
behavioral_data_clean_ANOVA <- ezANOVA(
  behavioral_data_clean,
  dv = .(ISIms),
  wid = .(p),
  within = .(targetIndex),
  return_aov = TRUE,
  detailed = TRUE
)

# Change effect size to partial eta squared (pes)
behavioral_data_clean_ANOVA <- aovEffectSize(behavioral_data_clean_ANOVA, 
                                             effectSize = "pes")

# Create a nicely formatted table for results
aovDispTable(behavioral_data_clean_ANOVA)

```

## Post-Hoc Tests - Tukey's HSD
```{r}
library(nlme) # Add to library if we end up using this approach.
library(multcomp) # For glht pos-hoc tests

# Create a new model using lme.
planned_comparisons <- lme(ISIms ~ targetIndex, 
                           random = ~1|p/targetIndex, # Random effects resulting from participants 
                           data = behavioral_data_clean, 
                           method = "ML") # Maximum likelihood

# Post-hoc tests (Tukey).
post_hocs <- glht(planned_comparisons, linfct = mcp(targetIndex = "Tukey"))
summary(post_hocs)

```
